---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
data <- read.csv("activity.csv", na.string = TRUE)
```



## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(dplyr)
library(ggplot2)
library(data.table)

data <- read.csv("activity.csv")
good <- complete.cases(data) #Removing NAs
data <- data[good, ]
data$steps <- as.integer(data$steps)
data <- subset(data, steps > 0)

data$date <- as.Date(as.character(data$date), format = "%Y-%m-%d")
sum_steps <- data %>%
        group_by(date) %>%
        summarise_at(vars("steps"), sum)
ggplot(data = sum_steps, aes(x = date, y = steps)) + geom_bar(stat = "identity")

mean_steps <- data %>%
        group_by(date) %>%
        summarise_at(vars("steps"), mean)

median_steps <- data %>%
        group_by(date) %>%
        summarise_at(vars("steps"), median)

colnames(mean_steps) <- c("Date", "Mean.Steps")
colnames(median_steps) <- c("Date", "Median.Steps")
merge(mean_steps, median_steps)
```



## What is the average daily activity pattern?
```{r, echo=TRUE}
library(dplyr)
library(ggplot2)
library(data.table)

data <- read.csv("activity.csv")
good <- complete.cases(data) #Removing NAs
data <- data[good, ]
data$steps <- as.integer(data$steps)
data <- subset(data, steps > 0)

mean_steps <- data %>%
        group_by(interval) %>%
        summarise_at(vars("steps"), mean)

g <- ggplot(data = mean_steps, aes(x = interval, y = steps)) + 
  geom_line( color = "blue") + 
  geom_point(color = "black")
g + labs(title = "Average Daily Activity Pattern", 
         x = "5-minute Interval", 
         y = "Average Number of Steps Taken") +
  theme(plot.title = element_text(hjust = 0.5))
 
max_mean <- as.character(max(mean_steps$steps))
max <- as.list(mean_steps[grepl(max_mean, mean_steps$steps), ])
max_mean <- max$steps
max_interval <- max$interval
```


The 5-minute interval which contains the maximum number of steps on average 
across all the days in the dataset is on the 835-minute mark with a value of 352.4839

## Imputing missing values
```{r, echo=TRUE}
library(dplyr)
data <- read.csv("activity.csv")
na_values <- sum(is.na(data$steps))

data$date <- as.Date(as.character(data$date), format = "%Y-%m-%d")
mean_steps <- data %>%
        group_by(date) %>%
        summarise_at(vars("steps"), mean)

data[is.na(data)] <- mean(data$steps, na.rm = TRUE)

ggplot(data = data, aes(x = date, y = steps)) + geom_bar(stat = "identity")

mean_steps <- data %>%
        group_by(date) %>%
        summarise_at(vars("steps"), mean)

median_steps <- data %>%
        group_by(date) %>%
        summarise_at(vars("steps"), median)

colnames(mean_steps) <- c("Date", "Mean.Steps")
colnames(median_steps) <- c("Date", "Median.Steps")
merge(mean_steps, median_steps)
```

The values when imputing the missing values had a drastic change and overall the 
mean and the median of the steps decreased compare when the missing values are not included in the computation


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
data <- read.csv("activity.csv")
data$day <- weekdays(as.Date(as.character(data$date), format = "%Y-%m-%d"))
#data$date <- factor(data$date, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
data[is.na(data)] <- 0
good <- complete.cases(data) #Removing NAs
data <- data[good, ]
data$steps <- as.integer(data$steps)
data <- subset(data, steps > 0)

data$wday <- ifelse(data$day == "Monday", "Weekday",
                    ifelse(data$day == "Tuesday", "Weekday",
                           ifelse(data$day == "Wednesday", "Weekday",
                                  ifelse(data$day == "Thursday", "Weekday",
                                         ifelse(data$day == "Friday", "Weekday",
                                                ifelse(data$day == "Saturday", "Weekend",
                                                       ifelse(data$day == "Sunday", "Weekend", 
                                                              NA )))))))
data$wday <- as.factor(data$wday)    

mean_steps <- data %>%
        group_by(interval, wday) %>%
        summarise_at(vars("steps"), mean)


ggplot(data = mean_steps, aes(x = interval, y = steps)) + 
  geom_line(stat = "identity", color = "blue") + 
  facet_grid(rows = vars(wday))
```

